---
title: "Practical Machine Learning Assignment"
output: html_document
---

##1. Start with Setting Working Directory and loading of library

```{r}
setwd("D:/DataScience/Coursera/PracticalMachineLearning/PracticalMLAssignment")
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
```

##2. Set seed to 12345
```{r}
set.seed(12345)
```


##3. To load Data from Source
```{r}
trainURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
```

```{r}
testURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

```{r}
training <- read.csv(url(trainURL),na.strings=c("NA", "#DIV/0!",""))
testing <- read.csv(url(testURL), na.strings = c("NA", "#DIV/0!", ""))
```


##4. Partition 60% of Data to Training Data
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list = FALSE)
myTraining <- training[inTrain,]; myTesting <- training[-inTrain,]
dim(myTraining); dim(myTesting)
```


##5. Cleaning and tranforming data
```{r}
myDataNZV <- nearZeroVar(myTraining, saveMetrics = TRUE)
```

```{r}
myNZVvars <- names(myTraining) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")
myTraining <- myTraining[!myNZVvars]
dim(myTraining)
```

```{r}
myTraining <- myTraining[c(-1)]
```

```{r}
trainingV3 <- myTraining
for(i in 1:length(myTraining)) {
    if(sum(is.na(myTraining[,i]))/nrow(myTraining) >= .6) {
        for(j in 1:length(trainingV3)){
            if (length(grep(names(myTraining[i]), names(trainingV3)[j]))==1){
                trainingV3 <- trainingV3[,-j]
            }
        }
    }
}
dim(trainingV3)
```

```{r}
myTraining <- trainingV3
rm(trainingV3)
```

```{r}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58])
myTesting <- myTesting[clean1]
testing <- testing[clean2]

dim(myTesting)
```

```{r}
dim(testing)
```

```{r}
for(i in 1:length(testing)){
    for(j in 1:length(myTraining)){
        if(length(grep(names(myTraining[i]), names(testing)[j]))==1){
            class(testing[j]) <- class(myTraining[i])
        }
    }
}
testing <- rbind(myTraining[2, -58], testing)
testing <- testing[-1,]
```


##6. First part of prediction is using Machine Learning algorithm: Decision Tree
```{r}
modFitA1 <- rpart(classe ~ ., data = myTraining, method = "class")
```

```{r}
fancyRpartPlot(modFitA1)
```

```{r}
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
```

```{r}
confusionMatrix(predictionsA1, myTesting$classe)
```


##7. Second part of predictin is using Machine Learning algorithm: Random Forests
```{r}
modFitB1 <- randomForest(classe ~., data = myTraining)
```

```{r}
predictionsB1 <- predict(modFitB1, myTesting, type = "class") 
```

```{r}
confusionMatrix(predictionsB1, myTesting$classe)
```

Comparing Decision Tree and Random Forests, Random Forests yielded better results.

##8. As a result, using Random Forests to predict Test dataset.
```{r}
predictionsB2 <- predict(modFitB1, testing, type = "class")
```

Generate files with predictions to submit for assignment Quiz

```{r}
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
    }
}

pml_write_files(predictionsB2)
```